### 2021-02-13 paper.tex
initial latex compile

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf,10pt]{acmart}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{End-user Web Scraping For Customization}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

\author{Kapaya Katongo}
\affiliation{%
  \institution{MIT CSAIL}
  \city{Cambridge, MA}
  \country{USA}
}
\email{kkatongo@mit.edu}

\author{Geoffrey Litt}
\affiliation{%
  \institution{MIT CSAIL}
  \city{Cambridge, MA}
  \country{USA}
}
\email{glitt@mit.edu}

\author{Daniel Jackson}
\affiliation{%
  \institution{MIT CSAIL}
  \city{Cambridge, MA}
  \country{USA}
}
\email{dnj@csail.mit.edu}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
% \renewcommand{\shortauthors}{Trovato and Tobin, et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  Websites are malleable: users can install browser extensions and run
  arbitrary Javascript in the developer console to change them. However,
  this malleability is only accessible to programmers with knowledge of
  HTML, CSS, Javascript and the DOM. To broaden access to customization,
  our prior work developed an approach that empowers end-users to
  customize websites without traditional programming via a browser
  extension called Wildcard. Wildcard's customizations are powered by
  web scraping adapters which are currently written in Javascript by
  programmers. This means that end-users can only customize a website if
  a programmer has written an adapter for it. Furthermore, end-users do
  not have the ability to extend adapters in order to perform new
  customizations or repair adapters to fix broken customizations.

  In this paper, we extend Wildcard with a new system for \emph{end-user
  web scraping for customization} which enables end-users to create,
  extend and repair adapters by demonstration. We describe three design
  principles that guided our system's development and are applicable to
  other end-user web scraping and customization systems: (a) users
  should be able to scrape data and use it in a single, unified
  environment, (b) users should be able to extend and repair the
  programs that scrape data via demonstration and (c) users should
  receive live feedback during their demonstrations.

  We have successfully used our system to create, extend and repair
  adapters by demonstration on a variety of websites. We provide example
  usage scenarios that showcase each of our design principles.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
%% From HERE
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011066.10011069</concept_id>
<concept_desc>Software and its engineering~Integrated and visual development environments</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Integrated and visual development environments}
% To HERE

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{end-user programming, software customization, web browser extensions}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
%\begin{teaserfigure}
%  \includegraphics[width=\textwidth]{sampleteaser}
%  \caption{Seattle Mariners at Spring Training, 2010.}
%  \Description{Enjoying the baseball game from the third-base
%  seats. Ichiro Suzuki preparing to bat.}
%  \label{fig:teaser}
%\end{teaserfigure}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\hypertarget{sec:introduction}{%
\section{Introduction}\label{sec:introduction}}

Many websites on the internet do not meet the exact needs of all of
their users. End-user web customization systems like Chickenfoot
\citep{bolin2005}, Thresher \citep{hogue2005}, Sifter \citep{huynh2006}
and Vegemite {[}lin2009{]} help users tweak and adapt websites to fit
their unique requirements, ranging from reorganizing or annotating
content on the website to automating common tasks. In our prior work, we
presented Wildcard \citep{litt2020a}, a customization system which
enables end-users to customize websites through direct manipulation. It
does this by augmenting websites with a table view that shows their
underlying structured data. The table is bidirectionally synchronized
with the original website, so end-users can easily customize the website
by interacting with the table, including sorting and filtering data,
adding annotations, and running computations in a spreadsheet formula
language.

Wildcard has a key limitation. In order to enable end-users to customize
a website, a Javascript programmer first needs to code an adapter that
specifies how to scrape the website content and set up a bidirectional
synchronization with Wildcard's table view. Take Alice, an end-user who
uses Wildcard to customize her experience on Google Scholar as shown
above. Wildcard's sorting customization gives her the power to sort
publications by title which the website does not natively allow. She
thinks this will be useful on Weather.com to sort the ten day forecast
by the weather descriptions so that she can quickly find all the sunny
days. Unfortunately, a programmer has not coded an adapter for
Weather.com so Alice is unable to customize it. Additionally, if the
adapter for Google Scholar stops functioning when the website changes,
Alice has no recourse to repair it on her own.

In this paper, we describe an addition to Wildcard: a system that
enables end-users to create, extend and repair website adapters by
demonstration within the browser. Using this scraping system, an
end-user can perform end-to-end web customizations using Wildcard on
arbitrary websites, without ever needing to code an adapter. Through a
series of examples, we show that end-users can utilize our system to
successfully create Wildcard adapters on a variety of websites via
demonstration (Section~\ref{sec:demos}). We also describe key aspects of
our system and how web scraping for customization reveals a constraint
that simplifies the \emph{wrapper induction} \citep{kushmerick2000} task
used to generalize user demonstrations
(Section~\ref{sec:implementation}).

Our key contribution is a set of three design principles that guided the
development of our system, which also offer insights that might be
applied to other end-user web scraping and customization tools
(Section~\ref{sec:design-principles}):

\begin{itemize}
\tightlist
\item
  \textbf{Unified Environment}: Users should be able to scrape data and
  interact with the scraped data in a single, unified environment. This
  minimizes the barrier to fluidly switching back and forth between the
  two tasks, rather than treating them as entirely independent tasks.
\item
  \textbf{Editing By Demonstration}: Users should be able to not only
  create programs for scraping data by demonstration, but also extend
  and repair the programs by demonstration. This enables users to build
  on other user's work, and is especially important in the context of
  web scraping since programs that scrape data sometimes break as the
  underlying website changes.
\item
  \textbf{Live Programming}: Users should receive live feedback as they
  perform demonstrations. The system should indicate how it is
  generalizing from the user's example and what the resulting data will
  look like, so that the user can adjust their demonstrations on the fly
  and quickly arrive at the desired result.
\end{itemize}

Finally, we share our broader vision for web scraping for customization,
and some opportunities for future work, including a proposal for how
Wildcard's spreadsheet-like formula language might augment
demonstrations to provide end-users with more expressiveness in the web
scraping process (Section~\ref{sec:conclusion}).

\hypertarget{sec:demos}{%
\section{Motivating Examples}\label{sec:demos}}

In this section, we show how end-users can create, extend and repair
adapters for Wildcard via demonstration.

\hypertarget{creating-an-adapter}{%
\subsection{Creating An Adapter}\label{creating-an-adapter}}

Alice, our end-user from Section~\ref{sec:introduction}, wants to
customize her experience on Weather.com by sorting the ten-day forecast
based on the description of the weather on each day, allowing her to
easily view all the sunny days. She starts the adapter creation process
by clicking a context menu item within the Weather.com page, and hovers
over a data value she might like to scrape.

The system provides live feedback as Alice hovers, demonstrating the
\textbf{live programming} principle:

\begin{itemize}
\tightlist
\item
  The selected row of data is annotated in the page with a border, to
  indicate that she will be demonstrating values from within that row.
\item
  The selected column of data is highlighted in the page with a green
  background, to show how the system has generalized her demonstration
  across all the rows in the data.
\item
  A table view appears at the bottom of the screen, and displays how the
  values will appear in the data table.
\end{itemize}

Alice tries hovering over several other elements in the page, taking
advantage of the live feedback environment to decide what data would be
useful. After considering several options, she decides to save the date
field in the first column of the table, and commits the action by
clicking.

Next, she performs a similar process to fill the next column with the
weather descriptions. After filling both columns, she also tries
hovering over previously scraped data, and the toolbar at the top of the
page indicates which column corresponds to the previously scraped data.
Finally, she ends the adapter creation process and is able to
immediately sort the forecast by the weather description column, because
Wildcard provides a \textbf{unified environment} that combines both
scraping and customizing:

\hypertarget{extending-an-adapter}{%
\subsection{Extending An Adapter}\label{extending-an-adapter}}

Alice has previously used Wildcard to customize timeanddate.com. In
addition to sorting the list of holidays in a year by the day of the
week, she also wants to sort them by category so she can view all the
federal holidays together. Previously, she would have needed to find a
programmer to help her edit the adapter code to incorporate this
additional data, but using our system's support for \textbf{editing by
demonstration}, Alice can extend the adapter herself.

While viewing the website, she clicks the ``Edit Adapter'' button above
the Wildcard table to initiate the adapter editing process. As she
hovers over the currently scraped values, the columns they belong to are
highlighted. Finally, she clicks on ``Federal Holiday'' to add the new
column of data and saves the changes. Alice then proceeds to sort the
list by the type of holiday without the intervention of a programmer.

\hypertarget{repairing-an-adapter}{%
\subsection{Repairing An Adapter}\label{repairing-an-adapter}}

Alice next visits Google Scholar to look up references for her thesis
project. Unfortunately, the customization she had applied to sort
publications by their title (which is not natively supported by Google
Scholar) is no longer working. In fact, the column in the Wildcard table
that contained all the publication titles is empty, because the
website's internals changed and broke the adapter's scraping logic.
Alice can repair this on her own, again taking advantage of
\textbf{editing by demonstration}.

Alice initiates the editing process, and initially hovers over the
desired value to demonstrate the column she wants to scrape. However,
the \textbf{live programming} interface indicates to her that the values
would be populated into column D; instead, she wants the values to be
inserted into column A where they previously appeared. So, Alice clicks
on the symbol for column A to indicate that she wants to scrape the
values into that column. She then proceeds to re-apply her customization
to the website by sorting the publications by their title without the
intervention of a programmer.

\hypertarget{sec:implementation}{%
\section{System Implementation}\label{sec:implementation}}

We implemented our end-user web scraping system as an addition to the
Wildcard browser extension. We start by describing our implementations
of row and column generalization, live programming, and editing by
demonstration, and then discuss some of the current limitations of our
system.

\hypertarget{generalization-algorithms}{%
\subsection{Generalization Algorithms}\label{generalization-algorithms}}

In order to generate reusable scrapers from user demonstrations, our
system solves the \emph{wrapper induction} \citep{kushmerick2000} task:
generalizing from a small set of user-provided examples to a scraping
specification that will work on other parts of the website, and on
future versions of the website.

We take an approach similar to that used in other tools like Vegemite
\citep{lin2009} and Sifter \citep{huynh2006}:

\begin{itemize}
\tightlist
\item
  We generate a single \emph{row selector} for the website: a CSS
  selector that returns a set of DOM elements corresponding to
  individual rows of the table.
\item
  For each column in the table, we generate a \emph{column selector}, a
  CSS selector that returns the element containing the column value
  within that row.
\end{itemize}

One important difference is that our algorithm only accepts row elements
that have direct siblings with a similar structure. We refer to this as
the \emph{row-sibling} constraint. Later in this section, we describe
how the constraint provides a useful simplification of the wrapper
induction task and discuss the resulting limitations this puts on our
system in a section that follows.

When a user first demonstrates a column value, the generalization
algorithm is responsible for turning the demonstration into a row
selector that will correctly identify all the row elements in the
website and a column selector that will correctly identify the element
that contains the column value within a row element. During subsequent
demonstrations, the generalization algorithm uses the generated row
selector to find the row element that contains the column value and
generates a column selector which identifies the corresponding column
element.

At a high level, the generalization algorithm's challenge is to traverse
far enough up in the DOM tree from the demonstrated element to find the
element which corresponds to the row. We solve this using a heuristic;
the basic intuition is to find a large set of elements with similar
parallel structure. Consider the following sample HTML layout, which
displays a truncated table of superheroes, with each row containing some
nested structure:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{\textless{}body\textgreater{}}
    \KeywordTok{\textless{}h1\textgreater{}}\NormalTok{ Team Iron Man }\KeywordTok{\textless{}/h1\textgreater{}}
    \KeywordTok{\textless{}div}\OtherTok{ class=}\StringTok{’container’}\KeywordTok{\textgreater{}}
        \KeywordTok{\textless{}div}\OtherTok{ class=}\StringTok{\textquotesingle{}avenger\textquotesingle{}}\KeywordTok{\textgreater{}}
            \KeywordTok{\textless{}div}\OtherTok{ class=}\StringTok{\textquotesingle{}names\textquotesingle{}}\KeywordTok{\textgreater{}}
                \KeywordTok{\textless{}span}\OtherTok{ class=}\StringTok{\textquotesingle{}super\_hero\_name\textquotesingle{}}\KeywordTok{\textgreater{}}\NormalTok{ Iron Man }\KeywordTok{\textless{}/span\textgreater{}}
                \KeywordTok{\textless{}span}\OtherTok{ class=}\StringTok{\textquotesingle{}real\_name\textquotesingle{}}\KeywordTok{\textgreater{}}\NormalTok{ Tony Stark }\KeywordTok{\textless{}/span\textgreater{}}
            \KeywordTok{\textless{}/div\textgreater{}}
           \KeywordTok{\textless{}span}\OtherTok{ class=}\StringTok{’gender’}\KeywordTok{\textgreater{}}\NormalTok{ Male }\KeywordTok{\textless{}/span\textgreater{}}
        \KeywordTok{\textless{}/div\textgreater{}}
        \KeywordTok{\textless{}div}\OtherTok{ class=}\StringTok{\textquotesingle{}avenger\textquotesingle{}}\KeywordTok{\textgreater{}}
            \KeywordTok{\textless{}div}\OtherTok{ class=}\StringTok{\textquotesingle{}names\textquotesingle{}}\KeywordTok{\textgreater{}}
                \KeywordTok{\textless{}span}\OtherTok{ class=}\StringTok{\textquotesingle{}super\_hero\_name\textquotesingle{}}\KeywordTok{\textgreater{}}\NormalTok{  Black Widow }\KeywordTok{\textless{}/span\textgreater{}}
                \KeywordTok{\textless{}span}\OtherTok{ class=}\StringTok{\textquotesingle{}real\_name\textquotesingle{}}\KeywordTok{\textgreater{}}\NormalTok{ Natalia Romanoff }\KeywordTok{\textless{}/span\textgreater{}}
            \KeywordTok{\textless{}/div\textgreater{}}
            \KeywordTok{\textless{}span}\OtherTok{ class=}\StringTok{’gender’}\KeywordTok{\textgreater{}}\NormalTok{ Female }\KeywordTok{\textless{}/span\textgreater{}}
        \KeywordTok{\textless{}/div\textgreater{}}
\NormalTok{        ...}
    \KeywordTok{\textless{}/div\textgreater{}}
\KeywordTok{\textless{}/body\textgreater{}}
\end{Highlighting}
\end{Shaded}

The user performs a demonstration by clicking on the SPAN element
containing ``Tony Stark.'' Our algorithm traverses upwards from the
demonstrated element, considering each successive parent element as a
potential candidate for the row element. For each parent element \(n\),
the process is as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  compute a selector \(s\) that when executed on \(n\) only returns the
  demonstrated element
\item
  for each sibling \(m\) of \(n\), execute \(s\) on \(m\) and record
  whether the selector returned an element. Intuitively, if the selector
  returns an element, this suggests that the sibling \(m\) has some
  parallel structure to \(n\)
\item
  compute \(n_{siblings}\), the number of sibling elements of \(n\) for
  which \(s\) returned an element
\end{enumerate}

Notice how the row-sibling constraint simplifies the problem: row
candidates without siblings that return an element after \(s\) is
executed on them have \(n_{siblings}\) = 0, thus disqualifying them.
Furthermore, a candidate element \(n\) is only accepted as the row
element if the row selector associated with it only matches itself and
elements that are its direct siblings.

From the parent of the initial SPAN element, the algorithm moves up the
DOM tree until it reaches the BODY element. Then, it generates a
selector for each \(n\) and discards any \(n\) with a selector that
matches elements that are either not itself or a direct sibling.
Finally, the algorithm picks an \(n\) with the largest, positive
\(n_{siblings}\), preferring nodes lower in the tree as a tiebreaker.
\(s\) is used as the selector for the column.

Applying this to our example, if the user clicks on the SPAN element
containing ``Tony Stark,'' , the first parent element is the DIV with
the class \emph{names}This DIV has a sibling SPAN element, but the
sibling doesn't return an element when \emph{.super\_hero}, the selector
that identifies the initial SPAN in the DIV, is executed on it. As a
result, it has \(n_{siblings}\) = 0. At the level above that, we
consider the DIV with class \emph{avenger}, which has at least one
sibling that returns an element when the selector \emph{.super\_hero} is
executed on it. Finally, the level above that once again has no siblings
that return an element when the selector \emph{.super\_hero} is executed
on them. Therefore, our algorithm returns the DIV element and outputs
\emph{.avenger} as the row selector and \emph{.super\_hero} as the
column selector. These selectors are used to generate a DOM scraping
adapter which returns the DOM elements corresponding to a superhero data
row in the table.

\hypertarget{live-programming}{%
\subsection{Live Programming}\label{live-programming}}

Live programming is implemented by continually running the
generalization algorithm on the DOM element under the user's cursor,
reverting if the user hovers away and committing when the user clicks.
The generated row and column selectors are used to highlight all the
matching elements on the website and create an adapter. Highlighting all
the matching column elements on the website provides visual feedback
about the system's generalization to the user. Creating an adapter
enables the system to populate the table view and set up the
bidirectional synchronization. Because the table is populated and the
bidirectional synchronization is set up, users can customize as they
scrape.

\hypertarget{editing-by-demonstration}{%
\subsection{Editing By Demonstration}\label{editing-by-demonstration}}

Our system generates adapters with the row selector and the column
selectors used to scrape the data. The row selector is a CSS selector
that identifies all the row elements of the data and the column
selectors are CSS selectors that identify each column's column elements.

When the editing process is initiated, the adapter's row selector and
column selectors are used to highlight the previously scraped values on
the website. Furthermore, the generalization algorithm takes the
adapter's row selector and uses it as the basis to generate new column
selectors after each demonstration. When a new column is demonstrated,
our system appends the generated column selector to the list of column
selectors. This is how adapters are extended to create new columns. When
an existing column is demonstrated, our system replaces the column's
current column selector with the generated column selector. This is how
adapters are repaired to fix broken columns.

Extending and repairing adapters in this manner is feasible because
column selectors are independent of each other: changing one column's
selector does not affect another column's selector. This is not the case
for system's in which the output of demonstrations are dependent on each
other. For example, in a web automation sequence that involves clicking
on a button to open a menu and then entering text into the menu's text
input, the step that enters the text is not independent because it
depends on the step that clicks the button to open the menu.

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

The row-sibling constraint we mentioned earlier is important for the end
goal of customization because row elements that are not direct siblings
may not represent data on the website that should be related as part of
the same table by customizations such as sorting and filtering. Take the
following sample HTML layout showing two tables of superheros (Team Iron
Man and Team Captain America):

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{\textless{}body\textgreater{}}
  \KeywordTok{\textless{}h1\textgreater{}}\NormalTok{ Team Iron Man }\KeywordTok{\textless{}/h1\textgreater{}}
  \KeywordTok{\textless{}div}\OtherTok{ class=}\StringTok{’container’}\KeywordTok{\textgreater{}}
      \KeywordTok{\textless{}div}\OtherTok{ class=}\StringTok{\textquotesingle{}avenger\textquotesingle{}}\KeywordTok{\textgreater{}}
          \KeywordTok{\textless{}span}\OtherTok{ class=}\StringTok{\textquotesingle{}super\_hero\_name\textquotesingle{}}\KeywordTok{\textgreater{}}\NormalTok{ Iron Man }\KeywordTok{\textless{}/span\textgreater{}}
      \KeywordTok{\textless{}/div\textgreater{}}
      \KeywordTok{\textless{}div}\OtherTok{ class=}\StringTok{\textquotesingle{}avenger\textquotesingle{}}\KeywordTok{\textgreater{}}
          \KeywordTok{\textless{}span}\OtherTok{ class=}\StringTok{\textquotesingle{}super\_hero\_name\textquotesingle{}}\KeywordTok{\textgreater{}}\NormalTok{  Black Widow }\KeywordTok{\textless{}/span\textgreater{}}
      \KeywordTok{\textless{}/div\textgreater{}}
\NormalTok{      ...}
  \KeywordTok{\textless{}/div\textgreater{}}
  \KeywordTok{\textless{}h1\textgreater{}}\NormalTok{ Team Captain America }\KeywordTok{\textless{}/h1\textgreater{}}
  \KeywordTok{\textless{}div\textgreater{}}
      \KeywordTok{\textless{}div}\OtherTok{ class=}\StringTok{\textquotesingle{}avenger\textquotesingle{}}\KeywordTok{\textgreater{}}
          \KeywordTok{\textless{}span}\OtherTok{ class=}\StringTok{\textquotesingle{}super\_hero\_name\textquotesingle{}}\KeywordTok{\textgreater{}}\NormalTok{ Captain America }\KeywordTok{\textless{}/span\textgreater{}}
      \KeywordTok{\textless{}/div\textgreater{}}
      \KeywordTok{\textless{}div}\OtherTok{ class=}\StringTok{\textquotesingle{}avenger\textquotesingle{}}\KeywordTok{\textgreater{}}
          \KeywordTok{\textless{}span}\OtherTok{ class=}\StringTok{\textquotesingle{}super\_hero\_name\textquotesingle{}}\KeywordTok{\textgreater{}}\NormalTok{  Scarlet Witch }\KeywordTok{\textless{}/span\textgreater{}}
      \KeywordTok{\textless{}/div\textgreater{}}
\NormalTok{      ...}
  \KeywordTok{\textless{}/div\textgreater{}}
\KeywordTok{\textless{}/body\textgreater{}}
\end{Highlighting}
\end{Shaded}

Without the constraint that row elements have to be direct siblings, the
row generalization algorithm could determine the row selector to be
\emph{.avenger} because it matches the largest number of parallel
structures (has the largest \(n_{siblings}\)). While this may be the
correct result for the task of extraction, it is not for the task of
customization. The selector matches all rows across the two tables so
the sorting and filtering customizations could place rows in the
incorrect table and thereby distort what the tables represent. Because
of this, our system currently does not support such DOM structures but
we plan to explore the possibility of extracting multiple tables from a
website and joining them.

Another structure that our system cannot currently generalize over is
one in which column elements are not contained within a row element that
is a direct sibling of all the other row elements. Take the following
sample HTML layout showing a table of superheros:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{\textless{}body\textgreater{}}
  \KeywordTok{\textless{}h1}\OtherTok{ class=}\StringTok{\textquotesingle{}super\_hero\_name\textquotesingle{}}\KeywordTok{\textgreater{}}\NormalTok{ Iron Man }\KeywordTok{\textless{}/h1\textgreater{}}
  \KeywordTok{\textless{}span}\OtherTok{ class=}\StringTok{\textquotesingle{}real\_name\textquotesingle{}}\KeywordTok{\textgreater{}}\NormalTok{ Real Name: Tony Stark }\KeywordTok{\textless{}/span\textgreater{}}
  \KeywordTok{\textless{}h1}\OtherTok{ class=}\StringTok{\textquotesingle{}super\_hero\_name\textquotesingle{}}\KeywordTok{\textgreater{}}\NormalTok{ Captain America }\KeywordTok{\textless{}/h1\textgreater{}}
  \KeywordTok{\textless{}span}\OtherTok{ class=}\StringTok{\textquotesingle{}real\_name\textquotesingle{}}\KeywordTok{\textgreater{}}\NormalTok{ Real Name: Steven Rogers }\KeywordTok{\textless{}/span\textgreater{}}
  \KeywordTok{\textless{}h1}\OtherTok{ class=}\StringTok{\textquotesingle{}super\_hero\_name\textquotesingle{}}\KeywordTok{\textgreater{}}\NormalTok{ Black Widow }\KeywordTok{\textless{}/h1\textgreater{}}
  \KeywordTok{\textless{}span}\OtherTok{ class=}\StringTok{\textquotesingle{}real\_name\textquotesingle{}}\KeywordTok{\textgreater{}}\NormalTok{ Real Name: Natalia Romanoff }\KeywordTok{\textless{}/span\textgreater{}}
\NormalTok{  ...}
\KeywordTok{\textless{}/body\textgreater{}}
\end{Highlighting}
\end{Shaded}

This page contains one table of data in which rows are made up of an H1
tag (super hero name) and a SPAN tag (real name). For the task of
extraction, the DOM elements can be scraped using the shown classes to
output a table in which the values of the H1 tags form the first column
and the values of the SPAN tag form the second column. For the task of
customization, our system would need to know what the row boundaries
were in order for customizations such as sorting and filtering to work
as expected. In Wildcard, this can be done with a hand-coded adapter by
creating artificial row boundaries. In the above DOM structure, an
artificial row boundary could be created by representing a row as an H1
tag and the SPAN tag that immediately follows it. These artificial rows
would form a single unit which would not distort the website after a
sorting or filtering customization. We plan to explore how such
artificial row boundaries might be created via demonstration in order to
support these structures (which are not uncommon, and can be seen, for
example, in HackerNews).

\hypertarget{sec:design-principles}{%
\section{Design Principles}\label{sec:design-principles}}

Below, we discuss the three design principles underlying our work and
how they relate to the broader field of end-user web scraping and
customization.

\hypertarget{unified-environment}{%
\subsection{Unified Environment}\label{unified-environment}}

In the previous iteration of Wildcard, web scraping was an entirely
separate activity from customization. Programmers that wrote scraping
adapters would need to switch into an IDE to write code as part of
customizing a new website, making it a much less unified environment.

This type of divide between tasks appears in other domains. In data
science, workflows revolve between cleaning and using data but this
often happens in different environments. The creators of Wrex
\citep{drosos2020}, a programming-by-example system for data wrangling,
reported that ``although data scientists were aware of and appreciated
the productivity benefits of existing data wrangling tools, having to
leave their native notebook environment to perform wrangling limited the
usefulness of these tools.'' In web scraping, if a user comes across an
omission while working with data scraped from a website, they need to
switch from the environment in which they are using the data to the
environment in which they wrote their scraping code in order to edit and
re-run it. Sometimes, the divide is more subtle. Vegemite
\citep{lin2009}, a system for end-user programming of mashups, had
separate steps for scraping data from a website and augmenting that data
via automation tasks. During a user study, participants reported that
``it was confusing to use one technique to create the initial table, and
another technique to add information to a new column.''

In this work, we have combined scraping and customization into a unified
activity within a single environment. The goal is to minimize the
environment switch between \emph{extracting} the data and \emph{using}
the data. A user might start out by scraping some data, and then switch
to customizing the website using the results. Then, they might realize
they need more data to perform their desired task, at which point they
can easily augment the adapter by demonstrating new columns. All of
these tasks take place right in the browser, where the user was
initially already using the website. Instead of bringing the data to
another tool, we have brought a tool to the data. This directly relates
to the idea of ``in-place toolchains'' \citep{zotero-60} being an
integral part of end-user programming systems.

Of course, there is value in specialized tools: Wildcard has nowhere
near the full capabilities of spreadsheet software or databases.
Nevertheless, we believe a single, unified environment for scraping and
customization presents a significantly lower barrier to entry for
customization.

\hypertarget{editing-by-demonstration-1}{%
\subsection{Editing By Demonstration}\label{editing-by-demonstration-1}}

Many end-user web scraping and macro systems allow users to
\emph{create} programs by demonstration but do not offer a way to
\emph{edit} them by demonstration. In Rousillon \citep{chasins2018}, a
web scraping program created by demonstration can only be edited through
a high-level, block-based programming language called Helena
\citep{zotero-51} which does not support providing new demonstrations to
extend or repair the scraping code. In Vegemite \citep{lin2009}, a web
automation program created through demonstration can only be edited by
editing the text-based representation of the program.

In Wildcard, if a website's hand-coded adapter ceases to function
because the website changes, the table may lose its data which means
that an end-user's customizations may also cease to function.
Furthermore, end-users are powerless to extend the scraping adapter to
add columns to the table in order to perform new customizations. This
goes against MacLean et. al.'s vision of user-tailorable systems
\citep{maclean1990} that give users ``a feeling of ownership of the
system, to feel in control of changing the system and to understand what
can be changed.'' Providing an easy way for users to edit programs is
therefore fundamental to fully democratizing web customization.

Editing by demonstration makes end-users first-class citizens in the
customization ecosystem. Because users interact with the scraped data
directly in the context of the website, it is easy to initiate the
scraping system in editing mode: the scraping system is simply booted up
using metadata stored with the scraping adapter to the state when the
demonstration was completed. Users that have gone through the creation
process will immediately realize what to do in order to extend or repair
the adapter. Users that have not gone through the creation process might
have a harder time but we provide visual clues (such as highlighting the
row to perform demonstrations from with a green border) and live
programming (immediately preview the results of demonstrations) that
serve as guides.

As discussed in Section~\ref{sec:implementation}, editing by
demonstration in the web scraping domain is feasible because column
selectors are independent of each other. However, this is not the case
with row selectors because column selectors are dependent on them. Our
system therefore does not support editing rows but this an acceptable
limitation given our focus on extension and repair which only involve
column selectors.

\hypertarget{live-programming-1}{%
\subsection{Live Programming}\label{live-programming-1}}

In some end-user web scraping systems like Rousillon
\citep{chasins2018}, users only get \emph{full} feedback about the
program's execution (generalization and the scraped values) after
providing all the demonstrations. This means they cannot adjust their
demonstrations in response to the system's feedback as they demonstrate.

Our end-user web scraping system employs live programming techniques to
eliminate this edit-compile-debug cycle by running the generalization
algorithm and generating an adapter after each user demonstration. As we
showed in Section~\ref{sec:demos}, when a user demonstrates a value of a
column they wish to scrape, our system immediately shows how it has
generalized the user's demonstration across the other rows of the data
by highlighting the all relevant values. It also populates the table
with the scraped data based on the latest demonstration. The
highlighting and table population serve to give users a view of how
their demonstration has been generalized and what data will be available
in the table once scraped.

Many successful end-user programming systems such as spreadsheets and
SQL provide users with immediate results after entering commands. Our
live programming environment is particularly similar to that of
FlashProg \citep{mayer2015}, a framework that provides user interface
support for programming-by-demonstration systems like FlashExtract
\citep{le2014}, and relates to the idea that an important quality of
end-user programming is ``interaction with a living system''
\citep{zotero-60}.

Unlike text-based commands which are only valid when complete (e.g
\texttt{SELECT\ *\ FRO} versus \texttt{SELECT\ *\ FROM\ user\_table}),
the target of demonstration commands (the value of a DOM element under
the cursor) is the same during both hover and click (incomplete command
verus complete command). This allows us to take a small step further by
executing a command before a user completes it, thereby providing them
with a preview of the results on hover.

There are limits to this approach. Providing live feedback on websites
with a large number of DOM elements or complex CSS selectors can slow
down the generalization process, especially if a user is constantly
moving their cursor. Furthermore, many datasets are too large to preview
in the table in their entirety; the user might benefit more from the
live feedback if it could summarize large datasets. For example,
FlashProg provides a summary of the generalization through a color-coded
minimap next to the scrollbar of its extraction interface.

\hypertarget{sec:related-work}{%
\section{Related Work}\label{sec:related-work}}

End-user web scraping for customization relates to existing work in
end-user web scraping by a number of tools.

FlashProg \citep{mayer2015} is a framework that provides user interface
support for FlashExtract \citep{le2014}, a framework for data scraping
by examples. FlashProg's interface provides immediate visual feedback
about the generalization and scrapes the matched values in an output
tab. In addition, it has a program viewer tab that contains a high level
description of what the generated program is doing and provides a list
of alternative programs. Finally, it has a disambiguation tab that
utilizes conversational clarification to disambiguate programs, the
conservations with the user serving as inputs to generate better
programs. Though FlashProg has many desirable features we aim to
implement in future iterations, it does not offer a unified environment
within a browser for scraping and customizing websites.

Roussillon \citep{chasins2018} is a tool that enables end-users to
scrape distributed, hierarchical web data. Its interface does not
provide \emph{full} live feedback about its generalizations or the
values to be scrapped until all the demonstrations have been provided
and the generated program has been run. If run on a website it has
encountered before, Roussillon makes all the previously determined
generalizations visible to the user by color-coding the values on the
website that belong to the same column. This is a desirable feature for
our system as users will not have to actively explore in order to
discover which values are available for scraping and how they are
related to each other. On the extension and repair front, Roussillon
presents the web scraping code generated by demonstration as an
editable, high-level, block-based language called Helena
\citep{zotero-51}. While Helena can be used to perform more complex
editing tasks like adding control flow, it presents a change in the
model used for creation. Our system maintains the model used for
creation by allowing users to extend and repair web scraping code via
demonstration.

Vegemite \citep{lin2009} is a tool for end-user programming of mashups.
It has two interfaces: one for scraping values from a website and
another for creating scripts that operate on the scraped values. The web
scraping interface does not provide live feedback about the
generalization on hover but after a user clicks, the interface shows the
result of the system's generalization by highlighting the all matched
values. Furthermore, even though the interface also has a table, the
table is only populated with the scraped values after all the
demonstrations have been provided. The scripting interface utilizes
CoScripter \citep{leshed2008} which is used to record operations on the
scraped values for automation. For example, the scripting interface can
be used to demonstrate the task of copying an address in the table,
pasting it into a walk score calculator and pasting the result back into
the table. The script would then be generalized to all the rows and
re-run to fill in the remaining walk scores. CoScripter provides the
generated automation program as text-based commands, such as ``paste
address into `Walk Score' input,'' which can be edited after the program
is created via ``sloppy programming'' \citep{lin2009} techniques.
However, this editing does not extend to the web scraping interface used
for demonstrations.

Sifter \citep{huynh2006} is a tool that augments well structured
websites with advanced sorting and filtering functionality. Like
Wildcard, it uses web scraping to extract data from websites in order to
enable customizations (sorting and filtering). It performs the web
scraping automatically using a variety of heuristics and solicits
guidance from the user if this fails. While automatic web scraping seems
desirable, it is unclear how useful it is if the goal is customization.
Given a row with ten scrapable values, and therefore ten columns, would
a user prefer to simply demonstrate the value for the single column they
are interested in or un-demonstrate the nine values they are not
interested in? This is a question we can only answer after performing a
user study.

\hypertarget{sec:conclusion}{%
\section{Conclusion And Future Work}\label{sec:conclusion}}

In this paper, we presented our progress towards \emph{end-user web
scraping for customization}, to empower end-users in Wildcard's
ecosystem to create, extend and repair scraping adapters. There are
several outstanding issues and open questions we hope to address in
future work.

Like existing approaches, web scraping in our current implementation is
limited to what can be demonstrated. This is problematic if users want
to scrape the URL associated with a link element, which is not visible,
or only scrape a substring of a value. To solve this, we plan to harness
Wildcard's formula language. End-users will be able to use formulas
targeted at web scraping to access DOM element properties and
attributes. For example, a formula like
\texttt{=GetAttribute(link\_column,\ ‘href’)} could be used to scrape
URLs of link elements. End-users will also be able to use formulas
targeted at data processing. For example, a formula like
\texttt{=GetSubstring(name\_column,\ 1,\ 2)} could be used to scrape
substrings of text values. Such web scraping and data processing
formulas will give end-users some of the power available to programmers
that write web scraping code in Javascript which supports a wide variety
of DOM access and processing ability.

To assess our design principles, we plan to carry out a broader
evaluation of our system through a user study. Furthermore, we plan to
incorporate the program viewer and disambiguation features of FlashProg
\citep{mayer2015} to give users more insight and control into the
generalization process.

Our end goal is to empower end-users to customize websites in an
intuitive and flexible way, and thus make the web more malleable for all
of its users.

% \printbibliography

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{references.bib}

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.